Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal Model to Beyond

### 2024-05-01 main.tex
:construction: Update library

\begin{section}{Introduction}
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that can serve on demand.
	This is empowered by the recent advances in \gls{ris}, a programmable metasurface that recycles and redistributes ambient electromagnetic waves for improved wireless performance.
	A typical \gls{ris} consists of numerous low-power sub-wavelength non-resonant scattering elements, whose response can be engineered in real-time to manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019}.
	% It not only experiences negligible noise and supports full-duplex transmissions, but also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	It enables low-noise, full-duplex operation, and also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	The most popular \gls{ris} research direction is \emph{joint beamforming} design with transceivers for a specific performance measure, which has attracted significant attention in wireless communication \cite{Wu2019,Guo2020,Liu2022}, sensing \cite{He2022,Luo2022,Hua2023}, and power transfer literature \cite{Wu2020a,Feng2022,Zhao2022}.
	While passive beamforming suffers severe attenuation from double fading, it usually offers a squared asymptotic behavior than active beamforming (e.g., second-order array gain and fourth-order harvested power \cite{Zhao2022}).
	On the other hand, \gls{ris} can also be used for \emph{backscatter modulation} by periodically switching its reflection pattern within channel coherence time.
	This creates a free-ride message stream (similar to index modulation \cite{Basar2017}) with dual benefits: integrating with the legacy transmitter for enhanced channel capacity \cite{Karasik2020,Basar2020,Ye2022} or serving as a dedicated source for low-power uplink communication \cite{Liang2020,Zhao2022a,Yang2024}.
	Different from above directions, \emph{channel shaping} exploits the \gls{ris} as a stand-alone device to modify the inherent properties of the propagation environment.
	This provides a ubiquitous wave scattering benchmark for different wireless applications and helps to decouple the \gls{ris}-transceiver design.
	Relevant shaping metrics can be classified into two categories:
	\begin{itemize}
		\item \emph{Singular value centric:} Closely related to the performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to numerical perturbations. The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}.
		\item \emph{Power centric:} The second-order statistics are less informative in \gls{mimo} but easier to analyze and optimize. The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} and leakage interference \cite{Santamaria2023a}.
	\end{itemize}

@@

	Fourth, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} achievable rate maximization problem with two beamforming solutions: a local-optimal approach via \gls{ao} and a low-complexity approach over channel shaping.
	The former iteratively updates active beamforming by eigenmode transmission and passive beamforming by geodesic \gls{rcg} until convergence.
	The latter suboptimally decouples the joint design into a channel power gain maximization subproblem and a conventional \gls{mimo} transmission subproblem, then propose a closed-form two-stage solution.
	We observe that \gls{bd}-\gls{ris} yields higher channel power and achievable rate than diagonal \gls{ris}, while the relative gains scale with the group size and \gls{mimo} dimensions.
	% especially with large group sizes and \gls{mimo} dimensions.
	Moreover, the rate difference between the optimal and shaping-inspired designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	% Interestingly, the rate difference between the optimal and low-complexity designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	Those results emphasize the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems and suggest channel shaping offers a promising path to decouple joint \gls{ris}-transceiver designs.
	% a promising channel shaping path towards practical \gls{ris} and transceiver design.
	% that channel shaping offers a promising path towards simplified and practical \gls{bd}-
	% It suggests that channel shaping offers a promising path towards simplified and practical \gls{bd}-\gls{ris} (and transceiver) designs.
	% via \gls{bd}-\gls{ris} offers a promising path towards simplified and practical design.
	% offers a promising path towards simplified and practical desi
	% It suggests channel shaping helps to decouple the joint \gls{ris}-transceiver design and offers a promising low-complexity solution.
	% This is the first work to discuss the implications of channel shaping on joint active-passive beamforming optimization.

>>>>

\begin{section}{Introduction}
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that can serve on demand.
	This is empowered by the recent advances in \gls{ris}, a programmable metasurface that recycles and redistributes ambient electromagnetic waves for improved wireless performance.
	A typical \gls{ris} consists of numerous low-power sub-wavelength non-resonant scattering elements, whose response can be engineered in real-time to manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019}.
	% It not only experiences negligible noise and supports full-duplex transmissions, but also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	It enables low-noise, full-duplex operation, and also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	The most popular \gls{ris} research direction is \emph{joint beamforming} design with transceivers for a specific performance measure, which has attracted significant attention in wireless communication \cite{Wu2019,Guo2020,Liu2022}, sensing \cite{He2022,Luo2022,Hua2023}, and power transfer literature \cite{Wu2020a,Feng2022,Zhao2022}.
	While passive beamforming suffers severe attenuation from double fading, it usually offers a squared asymptotic behavior than active beamforming (e.g., second-order array gain and fourth-order harvested power \cite{Zhao2022}).
	On the other hand, \gls{ris} can also be used for \emph{backscatter modulation} by periodically switching its reflection pattern within channel coherence time.
	This creates a free-ride message stream (similar to index modulation \cite{Basar2017}) with dual benefits: integrating with the legacy transmitter for enhanced channel capacity \cite{Karasik2020,Basar2020,Ye2022} or serving as a dedicated source for low-power uplink communication \cite{Liang2020,Zhao2024,Yang2024}.
	Different from above directions, \emph{channel shaping} exploits the \gls{ris} as a stand-alone device to modify the inherent properties of the propagation environment.
	This provides a ubiquitous wave scattering benchmark for different wireless applications and helps to decouple the \gls{ris}-transceiver design.
	Relevant shaping metrics can be classified into two categories:
	\begin{itemize}
		\item \emph{Singular value centric:} Closely related to the performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to numerical perturbations. The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}.
		\item \emph{Power centric:} The second-order statistics are less informative in \gls{mimo} but easier to analyze and optimize. The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} and leakage interference \cite{Santamaria2023a}.
	\end{itemize}

@@

	Fourth, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} \gls{pc} achievable rate maximization problem with two beamforming solutions: a local-optimal approach via \gls{ao} and a low-complexity approach over channel shaping.
	The former iteratively updates active beamforming by eigenmode transmission and passive beamforming by geodesic \gls{rcg} until convergence.
	The latter suboptimally decouples the joint design into a channel power gain maximization subproblem and a conventional \gls{mimo} transmission subproblem, then propose a closed-form two-stage solution.
	We observe that \gls{bd}-\gls{ris} yields higher channel power and achievable rate than diagonal \gls{ris}, while the relative gains scale with the group size and \gls{mimo} dimensions.
	% especially with large group sizes and \gls{mimo} dimensions.
	Moreover, the rate difference between the optimal and shaping-inspired designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	% Interestingly, the rate difference between the optimal and low-complexity designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	Those results emphasize the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems and suggest channel shaping offers a promising path to decouple joint \gls{ris}-transceiver designs.
	% a promising channel shaping path towards practical \gls{ris} and transceiver design.
	% that channel shaping offers a promising path towards simplified and practical \gls{bd}-
	% It suggests that channel shaping offers a promising path towards simplified and practical \gls{bd}-\gls{ris} (and transceiver) designs.
	% via \gls{bd}-\gls{ris} offers a promising path towards simplified and practical design.
	% offers a promising path towards simplified and practical desi
	% It suggests channel shaping helps to decouple the joint \gls{ris}-transceiver design and offers a promising low-complexity solution.
	% This is the first work to discuss the implications of channel shaping on joint active-passive beamforming optimization.
